### 2022-04-20 manuscript_covid19dashboard.tex
Updated the paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Interactive dashboard to monitor the COVID-19 outbreak},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Interactive dashboard to monitor the COVID-19 outbreak}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\textbf{Thiyanga S. Talagala}\footnote{Corresponding author, email:
  \href{mailto:ttalagala@gmail.com}{\nolinkurl{ttalagala@gmail.com}}}

Department of Statistics

Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO10230

\hspace{3cm}

\textbf{Randi Shashikala}

Department of Statistics, Faculty of Applied Sciences

University of Sri Jayewardenepura, Sri Lanka, CO 10230

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

As of September 20th, 2021, 221 countries and territories are infected
by the COVID-19 worldwide pandemic. Dashboards are the most often used
visualization method for visualizing COVID-19 data and informing the
public. The main objective of this study is to identify different
features, visualization methods \& improvements that should be occurred
by exploring the existing dashboards and to develop a dashbored to
visualize COVID-19 outbreak in Sri Lanka. We explored 15 different
dashboards. The most commonly used visualization methods in dashboard
development are bar charts, line charts, and interactive maps.
Dashboards that fit on a single screen are preferable than others.

\textbf{Keywords:} Visualization, multiple time series, heatmap,
COVID-19 vaccine, flexdashboard

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The COVID-19 pandemic is a global coronavirus illness outbreak that
began in 2019 and is being caused by the Severe Acute Respiratory
Syndrome Coronavirus 2 (SARS-CoV-2) virus. In December 2019, the first
COVID-19-infected patient was discovered in Wuhan, China. According to
Worldmeter data, there were 229835231 confirmed cases, 206515718
recovered cases, 18605877 active cases, and 4713636 deaths worldwide as
of September 20, 2021. Currently, 99.5 percent of active patients are in
a mild state, while 0.5 percent are in a critical state.

Dashboards are one of the greatest visual interpretation methods for
tracking the COVID-19 pandemic's spread and communication. There are a
plethora of COVID-19 visualization dashboards that have been designed to
visualize the pandemic's global and local status. Different software can
be used to generate dashboards. We explored 15 dashboards designed to
visualize COVID-19 data in the global and country levels. First,
dashboards were compared to identify the various features, visualization
approaches, and enhancements that should be implemented. Next, we
developed an interactive dashboard to visualized COVID-19 outbreak in
Sri Lanka.

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

The 15 dashboards we used in the study are listed below (Table 1).

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\newpage

\hypertarget{results}{%
\section{Results}\label{results}}

The ``Sri Lanka COVID-19 Dashboard'' provides an overview of the
COVID-19 pandemic for Sri Lanka. This dashboard has eight panels as
follows.

\textbf{Table 01: Description of the panels}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.13}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.87}}@{}}
\toprule
\textbf{Name of the Panel} & \textbf{Description of the Panel} \\
\midrule
\endhead
Overview & Total count of COVID-19 confirmed, recovered, deaths, active
cases \& total vaccine doses administered.\hfill\break Provide an
overview of daily COVID-19 confirmed, recovered \& deaths by
plots.\hfill\break \\
Cases by Wave & There are three tabs in this panel.\hfill\break  * Total
COVID-19 confirmed cases - Cumulative count of COVID-19 confirmed cases
with specific dates\hfill\break * COVID-19 Cases Distribution by Wave -
Daily confirmed cases by wave\hfill\break * COVID-19 Deaths Distribution
by Wave - Daily deaths by wave\hfill\break \\
COVID-19 Patients Distribution & Provide an overview of confirmed cases
district wise distribution. There are four tabs in this
panel.\hfill\break * Total COVID-19 Patients Distribution in Sri Lanka -
Total confirmed counts for each district is represented by tree
map\hfill\break * Country Map - Total confirmed cases in each district
represented by Sri Lanka country map\hfill\break * Distribution of Daily
COVID-19 Patients for Last 30 Days - Visualize the daily confirmed cases
distribution by districts in last 30 days\hfill\break * By Applying
Min-Max Transformation - Visualize the details in the third tab by
applying min-max transformation for each district\hfill\break \\
Vaccination Details & Provide an overview of COVID-19 vaccination in Sri
Lanka. There are two tabs.\hfill\break * Total Vaccine Doses - Visualize
the administered vaccine doses as first dose only \& fully
vaccinated\hfill\break * Total Administered Doses by Vaccine Name -
Visualize the vaccination by vaccine names\hfill\break \\
Top 10 Countries & In this panel, compare the Sri Lanka confirmed \&
deaths with top 10 countries in the world (top 10 countries - The
countries which have been reported highest number of confirmed cases as
31st of August 2021).\hfill\break There are two
tabs.\hfill\break * Comparison of Cumulative Cases in Sri Lanka with Top
10 Countries - Compare the confirmed and deaths in Sri Lanka with top 10
countries by cumulative time series plots\hfill\break * Comparison of
Log of Cumulative Cases in Sri Lanka with Top 10 Countries - Compare the
confirmed and deaths in Sri Lanka with top 10 countries by log
cumulative time series plots\hfill\break (The data has been pulled from
WHO)\hfill\break \\
Global Comparison & Compare the total confirmed \& deaths in Sri lanka
with top 10 countries in Global \& Asia. There are two
tabs.\hfill\break * Comparison of the sri Lanka with Top 10 Countries
Reporting the Most COVID-19 Cases in the World - Compare the total
confirmed \& deaths in Sri Lanka with top 10 countries in the world \&
compare the case fatality ratios\hfill\break * Comparison of the sri
Lanka with Top 10 Countries Reporting the Most COVID-19 Cases in the
Asia - Compare the total confirmed \& deaths in Sri Lanka with top 10
countries in the Asia \& compare the case fatality ratios\hfill\break \\
Global Map & Visualize the distribution of confirmed, recovered \&
deaths in the world by world map.\hfill\break \\
About & This panel contains the details about the dashboard. \\
\bottomrule
\end{longtable}

\hypertarget{discussion-and-further-research}{%
\section{Discussion and Further
research}\label{discussion-and-further-research}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}